Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en abstract

\maketitle
% modificar abstract para reflejar cambios en el paper
\begin{abstract} 
Decision making under uncertainty has been well studied when information is considered at the associative (probabilistic) level. The classical Theorems of von Neumann-Morgenstern and Savage provide a formal criterion for rational choice using purely associative information, which form the basis of many learning algorithms such as those used in Reinforcement Learning. In this work, we consider decision problems in which available actions and consequences are causally connected. We define a Causal Decision Problem and within this framework we state a previous result from J. Pearl, which relies on a known causal model and thus showing that it can be considered as a causal version of the classical von Neumann-Morgenstern Theorem. Furthermore, we consider the case when the causal mechanism that controls the environment is unknown to the decision maker, and propose and prove a causal version of Savage's Theorem. Then, we describe two applications for which these theorems provide theoretical foundations: causal games and optimal action learning in causal environments. These results highlight the importance of causal models in decision making and the variety of potential applications.

\keywords{Causality \and Decision Making \and Game Theory}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}
\tableofcontents
\newpage
\section{Introduction}
\label{intro}
Causal reasoning is a constant element in our lives as it is in human nature to constantly ask \textit{why} \citep{spirtes2000causation,pearl2018why}. Looking for causes is an everyday task and, in fact, causal reasoning is to be found at the very core of our minds \citep{waldmann2013causal,danks2014unifying,neil2019causality}. It has been argued that the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} in order to engage with the world \citep{friston2010free,clark2015surfing}. 
Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{tversky1977causal, tversky1980causal, lagnado2007beyond, hagmayer2008causal, hagmayer2009decision, hagmayer2017causality}.

>>>>

\maketitle
% modificar abstract para reflejar cambios en el paper
\begin{abstract} 
Causal reasoning is a fundamental aspect of intelligent reasoning since it allows to consider interventions and counterfactuals. Decision making under uncertainty has been well studied when information is considered at the associative (probabilistic) level. The classical Theorems of von Neumann-Morgenstern and Savage provide a formal criterion for rational choice using purely associative information, which form the basis of many learning algorithms such as those used in Reinforcement Learning. In this work, we consider decision problems in which available actions and consequences are causally connected. We define a Causal Decision Problem and within this framework we state a previous result from J. Pearl, which relies on a known causal model and thus showing that it can be considered as a causal version of the classical von Neumann-Morgenstern Theorem. Furthermore, we consider the case when the causal mechanism that controls the environment is unknown to the decision maker, and propose and prove a causal version of Savage's Theorem. Then, we describe two applications for which these theorems provide theoretical foundations: causal games and optimal action learning in causal environments. These results highlight the importance of causal models in decision making and the variety of potential applications.

\keywords{Causality \and Decision Making \and Game Theory}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}
\tableofcontents
\newpage
\section{Introduction}
\label{intro}
Causal reasoning is a constant element in our lives as it is in human nature to constantly ask \textit{why} \citep{spirtes2000causation,pearl2018why}. Looking for causes is an everyday task and, in fact, causal reasoning is to be found at the very core of our minds \citep{tversky1977causal,waldmann2013causal,danks2014unifying,neil2019causality}. It has been argued that the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} in order to engage with the world \citep{friston2010free,clark2015surfing}. 
Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{ tversky1980causal, lagnado2007beyond, hagmayer2008causal, hagmayer2009decision, hagmayer2017causality}.
